Increased productivity using Atomify - a real-time LAMMPS visualizer

### 2018-06-05 ms.tex
Wrote more case study

\lstdefinelanguage{LAMMPS}{
    morekeywords={units,atom_style,lattice,region,create_box,create_atoms,pair_style,pair_coeff,neigh_modify,mass,velocity,fix,run},
    sensitive=false, % keywords are not case-sensitive
    morecomment=[l]{//}, % l is for line comment
    morecomment=[s]{/*}{*/}, % s is for start and end delimiter
    morestring=[b]" % defines that strings are enclosed in double quotes
}

@@

\begin{figure}
	\centering
	\includegraphics[width=0.5\textwidth]{figures/initial_configuration.pdf}
	\caption{
		Several snapshots of the process of creating the simulation in our case study.
		The wrongly placed cylinder is shown in a) whereas it is placed correctly in b).
    }
	\label{fig:initial_configuration}
\end{figure}

We now want to delete atoms so that we have a certain density. The volume of the inner cylinder is $\pi R^2 L$ which can be calculated in lammps with \code{variable V equal PI*v\_R*v\_R*lx}.
In these units, we have chosen that the atomic mass is 1 so number density $\rho$ equals the mass density $\rho_m$.
The command \code{delete\_atoms} can be used: \code{delete\_atoms porosity region-ID fraction seed}, where we need to specify the fraction of the existing atoms in that region we want to delete.
To count the number of atoms inside a region, we first create a group containing all the atoms in the region and then create a variable that counts them.
Assuming a density $\rho = 0.05$, we can calculate the expected number inside the cylinder as $N = V\rho$.
The atoms inside the region can now be deleted. These steps can be done in LAMMPS like this
\begin{lstlisting}[basicstyle=\tiny, frame = none, numbers=none, framexleftmargin=0pt, xleftmargin=-0.75cm, xrightmargin=0.0cm]
	group cylinder1 region cylinder1
	variable N_cyl equal count(cylinder1)
	variable V equal PI*v_R*v_R*lx
	variable rho equal 0.1
	variable N_wanted equal $V*${rho}
	variable delete_fraction equal (${N_cyl}-${N_wanted})/${N_cyl}
	delete_atoms porosity cylinder1 ${delete_fraction} 1234
\end{lstlisting}
and the result is shown in figure \ref{fig:initial_configuration}c.

The next step now is to create the larger cylinder, containing the red atoms in figure \ref{fig:cylinder_simulation} where regular time integration will take place.

>>>>

\lstdefinelanguage{LAMMPS}{
    morekeywords={units,atom_style,lattice,region,create_box,create_atoms,pair_style,pair_coeff,neigh_modify,mass,velocity,fix,run, group every, equal, count, porosity, EDGE},
    sensitive=false, % keywords are not case-sensitive
    morecomment=[l]{//}, % l is for line comment
    morecomment=[s]{/*}{*/}, % s is for start and end delimiter
    morestring=[b]" % defines that strings are enclosed in double quotes
}

@@

\begin{figure}
	\centering
	\includegraphics[width=0.5\textwidth]{figures/initial_configuration.pdf}
	\caption{
		Several snapshots of the process of creating the simulation in our case study.
		This shows how one typically generate a geometry step by step where the immediate
		feedback is very useful to quickly discover mistakes.
		The wrongly placed cylinder is shown in a) whereas it is placed correctly in b).
		c) shows how the system looks after deleting atoms to obtain $\rho = 0.01$. The next cylinder $C_2$ is shown in d),
		and all regions (placed in groups) are shown in e). Here we have uniquely defined the purple shell which should have a dissipative thermostat.
		In f), the list of all groups defined in this system is shown with colors matching those of e). All figures are rendered with Atomify.
    }
	\label{fig:initial_configuration}
\end{figure}
We now want to delete atoms so that we have a certain density. The volume of the inner cylinder is $\pi R^2 L$ which can be calculated in lammps with \code{variable V equal PI*v\_R*v\_R*lx}.
In these units, we have chosen that the atomic mass is 1 so number density $\rho$ equals the mass density $\rho_m$.
The command \code{delete\_atoms} can be used: \code{delete\_atoms porosity region-ID fraction seed}, where we need to specify the fraction of the existing atoms in that region we want to delete.
To count the number of atoms inside a region, we first create a group containing all the atoms in the region and then create a variable that counts them.
Assuming a density $\rho = 0.05$, we can calculate the expected number inside the cylinder as $N = V\rho$.
The atoms inside the region can now be deleted. These steps can be done in LAMMPS like this
\begin{lstlisting}[basicstyle=\tiny, frame = none, numbers=none, framexleftmargin=0pt, xleftmargin=-0.75cm, xrightmargin=0.0cm]
	group cylinder1 region cylinder1
	variable N_cyl equal count(cylinder1)
	variable V equal PI*v_R*v_R*lx
	variable rho equal 0.1
	variable N_wanted equal $V*${rho}
	variable delete_fraction equal (${N_cyl}-${N_wanted})/${N_cyl}
	delete_atoms porosity cylinder1 ${delete_fraction} 1234
\end{lstlisting}
and the result is shown in figure \ref{fig:initial_configuration}c.

The next step now is to create a larger cylinder, the one containing the red atoms in figure \ref{fig:cylinder_simulation}, where regular time integration will take place.
This cylinder should have a slightly bigger radius as stated above and can be created with another region command:
\begin{lstlisting}[basicstyle=\tiny, frame = none, numbers=none, framexleftmargin=0pt, xleftmargin=-0.75cm, xrightmargin=0.0cm]
	region cylinder2 cylinder x $c $c $(v_R+4) EDGE EDGE
\end{lstlisting}
The expected result as can be seen in figure \ref{fig:initial_configuration}d.
Now we have the inner cylinder with the gas atoms defined as a region and the next cylinder shell (including the gas atoms) defined as another region.
We now want to create another shell, the one with the dissipative thermostat.
All the atoms with the dissipative thermostat, the inner shell and the gas will be integrated with the \code{fix nve} integrator.
Therefore, we create a slightly larger cylinder which will be the one defining the group for \code{fix nve}.
We can also define the groups with these cylinders
\begin{lstlisting}[basicstyle=\tiny, frame = none, numbers=none, framexleftmargin=0pt, xleftmargin=-0.75cm, xrightmargin=0.0cm]
	region cylinder3 cylinder x $c $c $(v_R+7) EDGE EDGE
	group cylinder1 region cylinder1
	group cylinder2 region cylinder2
	group cylinder3 region cylinder3
	group nve region cylinder3
\end{lstlisting}
where we also have created the group \code{nve} for clarity.

The dissipative thermostat will be \code{fix langevin}\citep{schneider1978molecular}, but it should only be applied on the outer shell defined by $C_3 \setminus C_2$,
where $C_i$ is the set containing the atoms inside cylinder $i$.
We can obtain this with the command
\begin{lstlisting}[basicstyle=\tiny, frame = none, numbers=none, framexleftmargin=0pt, xleftmargin=-0.75cm, xrightmargin=0.0cm]
	group langevin subtract cylinder3 cylinder2
	group frozen subtract all cylinder3
	group gas dynamic all region cylinder1 every 1
\end{lstlisting}
which will contain only the outer shell. 
We here also created the group \code{frozen}, which is the outmost atoms, and the group \code{gas} which is dynamic since atoms may switch places and we want to only apply a force on the ones in the inner cylinder at any point in time.
The frozen atoms will not move at all and make sure that the system does not start moving due to momentum transfer from the gas particles through the solid.
All these groups are highlighted in figure \ref{fig:initial_configuration}e by hovering the box containing all the groups. in Atomify as shown in figure \ref{fig:initial_configuration}f.
The colors of the atoms matches the group color next to the group name in figure \ref{fig:initial_configuration}f, where the order of coloring is the same as they were defined in the script.
Therefore we see no green atoms (\code{cylinder1}) since all of these are in the groups \code{cylinder2}, \code{cylinder3}, \code{nve} and \code{gas}.
